---
title: "PS630 Lab: RDD and Matching"
author: "Haohan Chen"
date: "November 30, 2018"
output: pdf_document
---

**Disclaimer: ** I'm using lots of online resources for this tutorial...

```{r, include=FALSE}
# Install packages
pkgs = c("rdd", "MatchIt", "tidyverse")
to_install = pkgs[which(!pkgs %in% installed.packages()[, 1])]
lapply(to_install, install.packages)
```

## Regression Discontinuity


The paper:

Carpenter, Christopher, and Carlos Dobkin. "The effect of alcohol consumption on mortality: regression discontinuity evidence from the minimum drinking age." American Economic Journal: Applied Economics 1, no. 1 (2009): 164-82.

**Topic: ** The effect of alcohol consumption on mortality

### Software Setup

```{r}
#---------------
# Package
#---------------
library(rdd)

#---------------
# Package
#---------------
library(readstata13)
AEJfigs = read.dta13("AEJfigs.dta")

#---------------
# Data cleaning
#---------------
# All = all deaths
AEJfigs$age = AEJfigs$agecell - 21
AEJfigs$over21 = ifelse(AEJfigs$agecell >= 21,1,0)
```

### RD Model: Age cutpoint and Overall Mortality

#### Model Fit

```{r}
#---------------
# Fit RD model
#---------------
reg.1=RDestimate(all~agecell,data=AEJfigs,cutpoint = 21)
```

#### Check Assumptions

```{r}
plot(reg.1)
title(main="All Causes of Death", xlab="AGE",ylab="Mortality rate from all causes (per 100,000)")
```

#### Results

```{r}
summary(reg.1)
```

### Other DVs: Death 

#### Motor Vehicle Accident

```{r}
reg.2=RDestimate(mva~agecell,data=AEJfigs,cutpoint = 21)
plot(reg.2)
title(main="Motor Vehicle Accidents Death", xlab="AGE",ylab="Mortality rate from all causes (per 100,000)")
summary(reg.2)
```

#### Internal Cause of Death

```{r}
reg.3=RDestimate(internal~agecell,data=AEJfigs,cutpoint = 21)
plot(reg.3)
title(main="Internal Causes of Death", xlab="AGE",ylab="Mortality rate from all causes (per 100,000)")
summary(reg.3)
```

**Further reading: ** https://rpubs.com/cuborican/RDD



## Matching

Example: 

"Causal Effects in Non-Experimental Studies: Reevaluating the Evaluation of Training Programs," Journal of the American Statistical Association, Vol. 94, No. 448 (December 1999), pp. 1053-1062. 

**Topic: ** The effect on trainee earnings of an employment program

### Software Setup

```{r}
#---------------
# load packages
#---------------
library(MatchIt)
# Political scientists package
# https://cran.r-project.org/web/packages/MatchIt/MatchIt.pdf

# Alternative packages: Matching, designmatch
# https://cran.r-project.org/web/packages/Matching/Matching.pdf
# https://cran.r-project.org/web/packages/designmatch/designmatch.pdf

library(tidyverse)

#---------------
# load dataset
#---------------
data("lalonde")
names(lalonde)
```

### Difference in Mean

```{r}
# Table
lalonde %>%
  group_by(treat) %>%
  select(treat, re74, re75, re78) %>%
  summarise_all(funs(mean, sd))
```

```{r}
# Plot (without error)
lalonde %>%
  group_by(treat) %>%
  select(treat, re74, re75, re78) %>%
  summarise_all(funs(mean, lo = quantile(., 0.025), hi = quantile(., 0.975))) %>%
  gather(key, value, -treat) %>%
  mutate(year = as.integer(substr(key, 3, 4)), 
         stat = substr(key, 6, 10),
         treat = as.factor(treat)) %>%
  select(-key) %>%
  spread(stat, value) %>%
  ggplot(aes(x = year, y = mean, color = treat)) + geom_point() + geom_line()
  
# Plot (with variance)
lalonde %>%
  group_by(treat) %>%
  select(treat, re74, re75, re78) %>%
  summarise_all(funs(mean, lo = quantile(., 0.025), hi = quantile(., 0.975))) %>%
  gather(key, value, -treat) %>%
  mutate(year = as.integer(substr(key, 3, 4)), 
         stat = substr(key, 6, 10),
         treat = as.factor(treat)) %>%
  select(-key) %>%
  spread(stat, value) %>%
  ggplot(aes(x = year, y = mean, color = treat)) + geom_point() + geom_line() +
  geom_errorbar(aes(ymin = lo, ymax = hi), width=0.2)

```

### Two-sample t-test

```{r}
# Earning Growth
#----------------
t.test(re78 - re75 ~ treat, data = lalonde)
# Remember what we did?
```

### Checking Balance in Covariates

```{r}
# Get a sense of the imbalance
lalonde %>%
  group_by(treat) %>%
  select(age, educ, black, hispan, married, nodegree) %>%
  summarise_all(funs(mean)) %>%
  gather(key, value, -treat) %>%
  spread(treat, value) %>%
  setNames(c("Covariate", "Control", "Treated")) %>%
  mutate(`T - C` = Treated - Control)
```

```{r}
# A statistical summary of their difference
vars = c("age", "educ", "black", "hispan", "married", "nodegree")
ttests = apply(lalonde[, vars], 2, function(x) t.test(x ~ lalonde$treat))

# T statistics
sapply(ttests, function(x) x$statistic["t"])

# p value
sapply(ttests, function(x) x$p.value)
```

### Propensity Score Estimation

#### Fit PS Model

```{r}
m_ps = glm(treat ~ age + educ + black + hispan + married + nodegree, 
           family = binomial(link="logit"), data = lalonde)
summary(m_ps)
```

#### Predict PS Score

```{r}
?predict
ps = predict(m_ps)
lalonde_ps = lalonde %>% mutate(propensity_score = ps)
```

#### Check overlap

```{r}
lalonde_ps %>%
  ggplot(aes(x = propensity_score, fill = factor(treat))) +
  geom_density(alpha = 0.5)
```

#### Matching with the propensity score

Some more complication. Luckily, we have an one-stop solution!

#### One-Stop Solution

```{r}
m_matchit = 
  matchit(treat ~ age + educ + black + hispan + nodegree + married + re74 + re75, 
        data = lalonde, method = "nearest", distance = "logit")

summary(m_matchit)
```

**Further reading: ** https://sejdemyr.github.io/r-tutorials/statistics/tutorial8.html