---
title: "PS630 Lab 1 Code"
author: "[Your Name], Haohan Chen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

```{r install, eval=FALSE}
# Execute this chunk of code first
install.packages("wbstats")
install.packages("ggplot2")
```

## Download and Install R and RStudio

- Download R from here: https://cloud.r-project.org/
- Download RStudio from here: https://www.rstudio.com/products/rstudio/download/

## A Crush Course of R Data Analysis

- The World Bank DataBank: http://databank.worldbank.org/data/home.aspx
- Use APIs to access data
- Suppose we want to analyze GDP per capita in the world in the year of 2016

### Load the World Bank R package

```{r}
#------------------------------------------------
# Our first data analysis: the World Bank data
#------------------------------------------------

# Install package: An R package that downloads data from the World Bank API
  # install.packages("wbstats")

# Load the World Bank data library
  library(wbstats)
```

### Load data (specific to this package)

```{r}
# Search for the indicator of a variable of interest
  search_result <- wbsearch(pattern = "GDP per capita")
  
# View search results
  # View(search_result) # Equivalent to clicking the name on the Data Panel
  
# View the first 20 rows of the data in the console
  head(search_result, 20)

# Download data of interest
  data <- wb(country = "all", indicator = "NY.GDP.PCAP.PP.CD", startdate = 2016, enddate = 2016)
```

### Simple Data cleaning

```{r}
# View data: what it looks like
  # View(data)

# Statistics of the dataset
  class(data) # the type of data structure
  dim(data) # dimension
  names(data) # columns
  
# After viewing the data, we decide we don't need the first 47 rows
  data_s1 <- data[-c(1:47), ]
  # View(data_s1)
  
# After viewing the data, we decide we need only the "country" and "value" column
  names(data_s1)
  data_s2 <- data_s1[, c("iso3c", "country", "value")]

# Rename the second column to "GDP_PAP_2016", so that the column names are more informative
  names(data_s2)[3] <- "GDP_PAP_2016"
  # View(data_s2)

# Get the final cleaned dataset
  data_f_gdppc <- data_s2
  
# Remove the intermediate product
  rm(data, search_result, data_s1, data_s2)
```

### Get another variable: Corruption

- Suppose our research question is about the correlation between corruption and economic development

```{r}
# Search for the indicator of a variable of interest
  search_result <- wbsearch(pattern = "corruption")
  
# View search results
  # View(search_result) # Equivalent to clicking the name on the Data Panel

# Download data of interest: control of corruption
  data <- wb(country = "all", indicator = "CC.EST", startdate = 2016, enddate = 2016)

# Data cleaning: delete rows? No

# Data cleaning: delete columns? Yes
  data_s1 <- data[, c("country", "value")]

# Change column name
  names(data_s1)[2] <- "corruption_control_2016"
    
# Final dataset
  data_f_corruption <- data_s1  

# Remove intermediate product
  rm(data, search_result, data_s1)
```

### Merging the two variables

```{r}
# Now we would like to merge the two data by the name of the country
  data_f <- merge(data_f_gdppc, data_f_corruption, by = "country")
# See the final product
  # View(data_f)
```

### Plots

Let's take a break from boring numbers. Draw some nice plots to describe the data

#### Plots of a single variable

```{r}
#-----------
# Histogram
#-----------
  # Simplist
  hist(data_f$GDP_PAP_2016)

  # add titles, labels
  hist(data_f$GDP_PAP_2016, main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)", ylab = "Number of countries")

  # Change the number of breaks
  hist(data_f$GDP_PAP_2016, breaks = 5, main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)", ylab = "Number of countries")
  hist(data_f$GDP_PAP_2016, breaks = 10, main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)", ylab = "Number of countries")
  hist(data_f$GDP_PAP_2016, breaks = 20, main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)", ylab = "Number of countries")

  # Add a line indicating the mean?
  hist(data_f$GDP_PAP_2016, breaks = 10, main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)")
  mean(data_f$GDP_PAP_2016)
  abline(v = 19922.52)
  
  # Change the style of the abline
  hist(data_f$GDP_PAP_2016, breaks = 10, main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)")
  abline(v = 19922.52, lty = 2, col = "red", lwd = 2) #lty: line style: dotted, dashed etc. lwd: width of the line
  
#-----------
# Density
#-----------
  # Default
  plot(density(data_f$GDP_PAP_2016))
  
  # add titles, labels
  plot(density(data_f$GDP_PAP_2016), main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)")

  plot(density(log(data_f$GDP_PAP_2016)), main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)")
  
  
  # Add a line indicating the mean
  plot(density(data_f$GDP_PAP_2016), main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)")
  abline(v = 19922.52, lty = 2, col = "red", lwd = 2) #lty: line style: dotted, dashed etc. lwd: width of the line

#-----------
# Boxplot
#-----------
  boxplot(data_f$GDP_PAP_2016, main = "GDP per capita 2016", xlab = "GDP per capita, PPP (current international $)")
```

#### Exercise: Plot statistics of the corruption variable

```{r}

```

#### Plot two variables

```{r}
#--------------
# Scatter plot
#--------------
  # default
  plot(data_f$GDP_PAP_2016, data_f$corruption_control_2016)
  
  # add titles, labels
  plot(data_f$GDP_PAP_2016, data_f$corruption_control_2016, xlab = "GDP per capita, PPP (current international $)", ylab = "Corruption Control Index", main = "Development and Corruption")
  
  # Change point size, style
  plot(data_f$GDP_PAP_2016, data_f$corruption_control_2016, xlab = "GDP per capita, PPP (current international $)", ylab = "Corruption Control Index", main = "Development and Corruption", pch = 24) #see this: https://www.statmethods.net/advgraphs/images/points.png
```

### Descriptive statistics

- For GDP per capita, let's calculate
    - Mean
    - Variance
    - Standard deviation
    - Other summary statistics
- For the two variables, calculate
    - Covariance
    - Correlation

```{r}
#-----------------
# Single variable
#-----------------

# Mean
  mean(data_f$GDP_PAP_2016)
# Variance
  var(data_f$GDP_PAP_2016)
# Standard deviation
  sd(data_f$GDP_PAP_2016)
# Quantiles
  quantile(data_f$GDP_PAP_2016)
  quantile(data_f$GDP_PAP_2016, c(0, 0.10, 0.20, 0.30))

# Summary
  summary(data_f$GDP_PAP_2016)
  
#-----------------
# Two variables
#-----------------

# Covariance
  cov(data_f$GDP_PAP_2016, data_f$corruption_control_2016)
# Correlation
  cor(data_f$GDP_PAP_2016, data_f$corruption_control_2016)
```


### More: Some prettier plots

```{r}
#-------------------------------------------------------
# GGPLOT: Another package that will be our best friend
#-------------------------------------------------------

# Install package
  # install.packages("ggplot2")
# Load package
  library(ggplot2)

# Plot Corruption, GDP per capita relation with country names as label
  ggplot(data_f, aes(x = GDP_PAP_2016, y = corruption_control_2016, label = country)) + 
    geom_point() + geom_text(aes(vjust = 1)) +
    ggtitle("Development and Corruption") + xlab("GDP per capita, PPP (current international $)") + ylab("Corruption Control Index")

# Plot Corruption, GDP per capita relation with iso3c country code as label
  ggplot(data_f, aes(x = GDP_PAP_2016, y = corruption_control_2016, label = iso3c)) + 
    geom_point() + geom_text(aes(vjust = 1)) +
    ggtitle("Development and Corruption") + xlab("GDP per capita, PPP (current international $)") + ylab("Corruption Control Index")

# Plot corruption, GDP per capita relation with a trend line
  ggplot(data_f, aes(x = GDP_PAP_2016, y = corruption_control_2016, label = iso3c)) + 
    geom_point() + geom_text(aes(vjust = 1)) + geom_smooth() +
    ggtitle("Development and Corruption") + xlab("GDP per capita, PPP (current international $)") + ylab("Corruption Control Index")

# Plot corruption, GDP per capita relation with a trend line: linear
  ggplot(data_f, aes(x = GDP_PAP_2016, y = corruption_control_2016, label = iso3c)) + 
    geom_point() + geom_text(aes(vjust = 1)) + geom_smooth(method = "lm") +
    ggtitle("Development and Corruption") + xlab("GDP per capita, PPP (current international $)") + ylab("Corruption Control Index")
```