---
title: 'PS630 Lab: Interaction Effect'
author: "Haohan Chen"
date: "November 16, 2018"
output: html_document
---

## Load the Package

```{r, message = F, warning = F}
library(readstata13)
d = readstata13::read.dta13("TeachingRatings.dta")

library(tidyverse)
```

## Fit a Model

```{r}
m = lm(course_eval ~ beauty + female + age + minority + age:minority + onecredit + nnenglish, data = d)
summary(m)
```

## What is the model?

$$
course\_eval_i = \beta_0 + \beta_1 beauty_i + \beta_2 female_i + \beta_3 age_i + \beta_4 minority_i + \beta_5 age_i \times minority_i  + \beta_6 onecredit_i + \beta_7 nnenglish_i + \varepsilon_i
$$

With the above model fit

$$
\hat{course\_eval_i} = \hat{\beta}_0 + \hat{\beta}_1 beauty_i + \hat{\beta}_2 female_i + \hat{\beta}_3 age_i + \hat{\beta}_4 minority_i + \hat{\beta}_5 age_i \times minority_i  + \hat{\beta}_6 onecredit_i + \hat{\beta}_7 nnenglish_i
$$

## Interpretation

### Question 1a: What is the marginal effect of `beauty` on `course_eval`?

$$
\frac{\partial \hat{course\_eval}}{\partial beauty} = \hat{\beta}_1
$$

Note: `beauty` is a continuous variable

Answer: One unit increase in `beauty` is associated with $\hat{\beta}_1$  change in `course_eval`.

### Question 1b

Hypothetically, if `beauty` is a dummy variable (say, `beauty=1` means good-looking and `beauty=0` means otherwise), what is the marginal effect?

Answer: When `beauty` change from 0 to 1, `course_eval` change by $\hat{\beta}_1$. 

### Question 2: What is the marginal effect of `age` on `course_eval`

$$
\frac{\partial \hat{course\_eval}}{\partial age} = \hat{\beta}_3 + \hat{\beta}_5 \times minority
$$

Note: This is a *continuous* variable.

Answer: One unit increase in `age` is associated with $\hat{\beta}_3 + \hat{\beta}_5 \times minority$ units increase of `course_eval`

### Question 3: What is the marginal effect of `minority` on `course_eval`

$$
\frac{\partial \hat{course\_eval}}{\partial minority} = \hat{\beta}_4 + \hat{\beta}_5 \times age
$$

Note: This is a *dummy* variable which only takes values 0 and 1.

Answer: When `minority` change from 0 to 1, `course_eval` change by $\hat{\beta}_4 + \hat{\beta}_5 \times age$.


## Obtaining the Coefficitns: Estimates and Uncertainty

We can obtain $\beta$'s and their confidence intervals with the functions provided by the `lm` package. We can also do it in a Monte Carlo way. That is, we simulate a set of coefficients to account for its uncertainty.

**You need not understand the details of the Monte Carlo approach**. It is actually much more straightfoward that it sounds. Useful materials:

- Multivariate Normal Distribution: https://en.wikipedia.org/wiki/Multivariate_normal_distribution. This Explain what we do with the `mvrnorm` function.
- Central Limit Theorem: This explain why we want to use multivariate normal distribution to simulate the coefficients. https://en.wikipedia.org/wiki/Central_limit_theorem
- A not-so-technical overview of this method with examples in R: https://aosmith.rbind.io/2018/01/09/simulate-simulate-part1/
 
```{r}
# Getting the coefficients and confidence interval from the model directly
# Easier way. Black box, not flexible
coef(m)
confint(m)

# The Monte Carlo approach
###########################
beta_m = coef(m) # mean (estimated) coefficients
beta_vcov = vcov(m) # Variance-covariance matrix of the coefficients
# Simulate beta's
set.seed(1111) # Set a seed so that you replicate the result
beta_sim = MASS::mvrnorm(n = 1000, mu = beta_m, Sigma = beta_vcov)
```

Let's take a look at what we get

```{r}
head(beta_sim)
```

We can first try to obtain what we can get from the `lm` function (`confint`)

```{r}
# Compare with what you get from confint
###########################################
beta_sim %>%
  apply(2, function(x) quantile(x, c(0.025, 0.975))) %>%
  t()
confint(m)
```

```{r}
# Get the 3 critical quantiles all together (2.5%, 50%, 97.5%)
###############################################################

coef_CI = beta_sim %>%
  apply(2, function(x) quantile(x, c(0.025, 0.05, 0.5, 0.95, 0.975))) %>%
  t()
coef_CI
```

## Plot the coefficients

```{r}
# How to plot it with ggplot?
coef_CI %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  ggplot(aes(x = rowname, y = `50%`)) +
  geom_point() +
  geom_errorbar(aes(ymin = `2.5%`, ymax = `97.5%`), width = 0.2) +
  # geom_linerange(aes(ymin = `5%`, ymax = `95%`), size = 1) + # if you want to plot 90% CI additionally
  geom_hline(yintercept = 0, color = "red", linetype = 2) + 
  xlab("Variable Names") + ylab("Coefficients (95% Confidence Interval)") +
  coord_flip()

# Aside: want to adjust the order of coefficients?
# Make row names a factor whose levels you manually order
# Copy the above code

coef_CI %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  mutate(rowname = factor(rowname, levels = rev(c("age", "minority", "age:minority", "beauty", "female", "onecredit", "nnenglish", "(Intercept)")))) %>%
  # Note on rev -- to take care of the consequence of coord_flip
  ggplot(aes(x = rowname, y = `50%`)) +
  geom_point() +
  geom_errorbar(aes(ymin = `2.5%`, ymax = `97.5%`), width = 0.2) +
  geom_linerange(aes(ymin = `5%`, ymax = `95%`), size = 1) +
  geom_hline(yintercept = 0, color = "red", linetype = 2) + 
  xlab("Variable Names") + ylab("Coefficients (95% Confidence Interval)") +
  coord_flip()

# Play with the above code by commenting out some lines to see what change.
```

## Plotting the Marginal Effect of a Variable

Back to the business: How do we plot the interaction effect?? Let's set it straight, what are the variables of the x-axis and the y-axis?

### Marginal effect of a continuous variable that has no interaction effect

Example: We would like to plot the marginal effect of `beauty` on `course_eval`

- x-Axis: Value of ?? (anything, the marginal effect is constant conditional on any other independent variables)
- y-Axis: Value of $\frac{\partial \hat{course\_eval}}{\partial beauty} = \beta_1$

No interaction effect. This means the marginal effect of ``beauty` is constant. There isn't much value to visualize it (though you can!)


### Marginal effect of a dummy variable that interacts with a continuous variable

Example: We would like to plot the marginal effect of `minority` on `course_eval`

- x-Axis: Value of `age`
- y-Axis: Value of $\hat{\beta}_4 + \hat{\beta}_5 \times age$

```{r}
# Step 0: Does the marignal effect depend on any other variables?
# Yes. Age

# Step 1: Get summary statistics of the variables identified in the above step
summary(d$age)

# Step 2: Simulate a sequence of the variable of interest in its range
age_sim = seq(min(d$age), max(d$age), length.out=20)

# Step 3: Multiply the sequence with the simulated beta's
quantiles = c(0.025, 0.5, 0.975)

me_minority = apply(beta_sim[, c("age", "age:minority")], 1, function(x) x["age"] * 1 + x["age:minority"] * age_sim) %>%
  apply(1, function(x) quantile(x, quantiles)) %>%
  as.data.frame() %>%
  t() %>%
  as.data.frame() %>%
  mutate(variable = age_sim)

me_minority

me_minority %>% 
  ggplot(aes(x = variable, y = `50%`)) +
  geom_line() + geom_errorbar(aes(ymin = `2.5%`, ymax = `97.5%`), color = "darkgray") +
  geom_hline(yintercept = 0, color = "red", linetype = 2) +
  xlab("age") + ylab("Marginal Effect of Minority Status")

me_minority %>% 
  ggplot(aes(x = variable, y = `50%`)) +
  geom_line() + geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = 0.3) + 
  geom_hline(yintercept = 0, color = "red", linetype = 2) +
  xlab("age") + ylab("Marginal Effect of Minority Status")
```

### Marginal effect of a continuous variable that interacts with a dummy variable

Example: We would like to plot the marginal effect of `minority` on `course_eval`

- x-Axis: Value of `beauty`
- y-Axis: Value of $\hat{\beta}_4 + \hat{\beta}_5 \times age$

```{r}
# Step 0: Does the marignal effect depend on any other variables?
# Yes. minority

# Step 1: Get summary statistics of the variables identified in the above step
table(d$minority)

# Step 2: Simulate a sequence of the variable of interest in its range
minority_sim = c(0, 1)

# Step 3: Multiply the sequence with the simulated beta's
quantiles = c(0.025, 0.5, 0.975)

me_age = apply(beta_sim[, c("minority", "age:minority")], 1, function(x) x["minority"] * 1 + x["age:minority"] * minority_sim) %>%
  apply(1, function(x) quantile(x, quantiles)) %>%
  as.data.frame() %>%
  t() %>%
  as.data.frame() %>%
  mutate(variable = minority_sim)

# Plot: using the original code: Bad
# Problem: An informative plot for a variable interacting with a dummy is slightly different.
me_age %>% 
  ggplot(aes(x = variable, y = `50%`)) +
  geom_line() + geom_errorbar(aes(ymin = `2.5%`, ymax = `97.5%`), color = "darkgray") +
  geom_hline(yintercept = 0, color = "red", linetype = 2) +
  xlab("minority") + ylab("Marginal Effect of Age")

me_age %>% 
  mutate(variable = factor(variable)) %>%
  ggplot(aes(x = variable, y = `50%`)) +
  geom_point() +
  geom_line(aes(group = 1)) + geom_errorbar(aes(ymin = `2.5%`, ymax = `97.5%`), width = 0.1) +
  geom_hline(yintercept = 0, color = "red", linetype = 2) +
  xlab("minority") + ylab("Marginal Effect of Age")
# This also apply for interaction with categorical variables (more than two classes)
```
