---
title: "More `dplyr` and `tidyr`"
author: 
  - "Haohan Chen^[Political Science Department, Duke University. haohan.chen@duke.edu]"
date: "October 5, 2018"
fontsize: 11pt
geometry: margin=1cm
linkcolor: blue
output: 
  pdf_document:
    number_sections: false
    keep_tex: true
    fig_caption: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


# Download an example dataset

```{r}
# Just to show how the data is downloaded, in case you are curious.
# Data downloade from ICPSR: https://www.icpsr.umich.edu/icpsrweb/ICPSR/

# if (!"psData" %in% installed.packages()[, 1]){
#   install.packages("psData")
# }
# 
# library(psData)
# polity <- PolityGet()
# save(polity, file = "polity.Rdata")
```

# Load the data and libraries

```{r}
# Libraries
library(dplyr)
library(tidyr)

# Data
load("polity.Rdata")
polity <- as_tibble(polity)
```

# Review: What you have learned

- `select`, `filter`, `arrange`, `mutate`. What do they do?
- Exercise: Get the variables `country` and `polity` of the "Uganda" from 1990-2000, sort it form largest to smallest.

Note: Note that POLITY score captures political regime authority spectrum on a 21-pont scale ranging from -10 (hereditary monarchy) to +10 (consolidated democracy). The Polity scores can also be converted into regime categories in a suggested three part categorization of "autocracies" (-10 to -6), "anocracies" (-5 to +5 and three special values: -66, -77 and -88), and "democracies" (+6 to +10). Performance score from 0 to 100. The highest score reflects the best situation.

```{r}
polity %>% filter(country == "Brazil") %>% select(country, year, polity2) %>% arrange(-polity2)
```

# New

- group_by, summarise: Get the mean, maximum, minimum, median, standard deviation of `polity2` for each country from 1990-2000.
- slice: Get the first 10 rows
- reshape dataset: long <- wide: Get a wide-form dataset of 2000-2005, Each year as a column

```{r}
polity %>% 
  filter(year %in% c(2000:2005)) %>%
  select(country, year, polity2) %>% 
  spread(year, polity2)
```



# More information

- **Must read:** https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf
- Some tutorials with examples:
    - https://rpubs.com/bradleyboehmke/data_wrangling
    - http://garrettgman.github.io/tidying/
