---
title: "PS630 Lab6 Notes"
author: "Haohan Chen"
date: "October 19, 2018"
output: pdf_document
---

# Load the data

```{r}
load("lab6_data_exercise.Rdata")
# Describe the dataset
str(data_country)
str(data_party)
```

# The package

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)

# Convert data frmes to tibbles
data_country <- as_tibble(data_country)
data_party <- as_tibble(data_party)
```



# Review of Labs 4 and 5

0. What does the `%>%` command do?
1. Create a new dataset from `data_country`, with only variables `country_code` and `GDP_pcap_ppp`
2. With the above dataset, create a variable `log_GDP_pcap_ppp` that is the logarithm of `GDP_pcap_ppp`
3. With the above data, create a new dataset from `data_country`, with only observations (countries) that are in the `Advanced Capitalist` category (variable `country_category`)
4. With the above data, sort observations by `GDP_pcap_ppp`, (1) from high to low; (2) from low to high
5. With the above data (or data of 3), find top 5 countries in terms of GDP per capita

```{r}
# Your code here
# 1 (select)
data_country %>% select(country_code, GDP_pcap_ppp)
data_country2 <- data_country %>% select(country_code, GDP_pcap_ppp)

# 2 (mutate)
# 2. With the above dataset, create a variable `log_GDP_pcap_ppp` that is the logarithm of `GDP_pcap_ppp`
data_country %>% 
  mutate(log_GDP_pcap_ppp = log(GDP_pcap_ppp))

mutate(data_country, log_GDP_pcap_ppp = log(GDP_pcap_ppp))

# 3 (filter)
# 3. With the above data, create a new dataset from `data_country`, with only observations (countries) that are in the `Advanced Capitalist` category (variable `country_category`)

data_country %>% 
  mutate(log_GDP_pcap_ppp = log(GDP_pcap_ppp)) %>%
  filter(country_category == "Advanced Capitalist"
         & polity2_score > 9)


# 4 (arrange)
#4. With the above data, sort observations by `GDP_pcap_ppp`, (1) from high to low; (2) from low to high

data_country %>% arrange(wt = GDP_pcap_ppp)

data_country %>% arrange(wt = -GDP_pcap_ppp)



# 5 (top_n)
# 5. With the above data (or data of 3), find top 5 countries in terms of GDP per capita

data_country %>% top_n(5, wt = GDP_pcap_ppp)

data_country %>% top_n(5, wt = -GDP_pcap_ppp)


```

# New: Summarize dataset

- `group_by(data, x)`: Group data into rows with the same value of `x`.
- `summarise(data, avg = mean(x))`: Summarise data into single row of values
- `summarise_all(data, funs(mean, var, sd))`: Apply summary function to each column. Example here: mean, variance, standard deviation
- `summarise_at(data, vars(x, y), funs(mean, var, sd))`: Apply summary function indicated in `funs() ` to a set of variables indicated in `vars()`

```{r}
# Getting started
data_country %>% group_by(country_category) %>%
  summarise(avg_GDP_pcap_ppp = mean(GDP_pcap_ppp))

# Anything special? How to handle it?
data_country %>% group_by(country_category) %>%
  summarise(avg_GDP_pcap_ppp = mean(GDP_pcap_ppp, na.rm = T))


# Want more than one summary statistics
data_country  %>% group_by(country_category) %>% 
  summarise(GDP_pcap_ppp_avg = mean(GDP_pcap_ppp, na.rm = T),
            GDP_pcap_ppp_var = var(GDP_pcap_ppp, na.rm = T),
            GDP_pcap_ppp_sd = sd(GDP_pcap_ppp, na.rm = T),
            nrow = n())

# Get multiple summary statistics conveniently with summarise_all, summarise_at
data_country %>% select(country_category, GDP_pcap_ppp) %>%
  group_by(country_category) %>%
  summarise_all(funs(mean, var, sd))

data_country %>% select(country_category, GDP_pcap_ppp, polity2_score) %>%
  group_by(country_category) %>%
  summarise_all(
    funs(mean(., na.rm = T), var(., na.rm = T), sd(., na.rm = T)))

mean(c(1, 2, 3, 4, NA))
mean(c(1, 2, 3, 4, NA), na.rm = T)


data_country %>% group_by(country_category) %>% 
  summarise_at(vars(GDP_pcap_ppp, polity2_score), 
               funs(min, median, max, mean, var, sd))

#
data_country %>% group_by(country_category) %>%
  mutate(avg_GDP_pcap_ppp = mean(GDP_pcap_ppp)) %>%
  select(country_code, GDP_pcap_ppp, avg_GDP_pcap_ppp, country_category)

```

**Exercise:** Get minimum, maximum and median of variables GDP_pcap_ppp and polity2_score by `country_category`. Follow the codes above.

```{r}
# Your code here
```


**General advice: ** Keep the original, raw data; Make informative names; Handle missing values carefully.

# New: Merging two datasets

An critical data management tool. Not intellectually challenging, but tedious.

## The Basics

Useful functions merging datasets from `dplyr` include `left_join`, `right_join`, `full_join`, `inner_join`.

- `left_join(a, b, by = "x1")`: Join matching rows from b to a. (Keep all in `a`)
- `right_join(a, b, by = "x1")`: Join matching rows from a to b. (keep all in `b`)
- `full_join(a, b, by = "x1")`: Join data. Retain all values, all rows. (keep all -- either in `a` or `b`)
- `inner_join(a, b, by = "x1")`: Join data. Retain only rows in both sets. (only keep those in both `a` and `b`)

```{r}
# Left join
merge_left <- data_country %>% left_join(data_party, by = "country_code")
dim(merge_left)

# Right join
merge_right <- data_country %>% right_join(data_party, by = "country_code")
dim(merge_right)

# Full join
merge_full <- data_country %>% full_join(data_party, by = "country_code")
dim(merge_full)


# Inner join
merge_inner <- data_country %>% inner_join(data_party, by = "country_code")
dim(merge_inner)

```

**Aside 1:** Matching on multiple identifiers?
```{r}
data_country %>% left_join(data_party, by = c("country_name", "country_code"))
```

**Aside 2:** Matching on different variable names?

```{r}
# Different variable names
data_country %>% left_join(data_party, by = c("country_name" = "country_name", 
                                              "country_code" = "country_code"))

```

## Advanced: Checking what match and what do not, and debug

**Checking what do not match**: A command good for checking `anti_join`
- `anti_join(a, b, by = "x1")`: All rows in a that do not have a match in b.

```{r}
# What countries cannot be matched with party info?
data_country %>% anti_join(data_party, by = "country_code")
data_party %>% anti_join(data_country, by = "country_code")


# Exercise: What parties cannot be matched with country info?

```




## General Advice

Be extremely careful in this step. It can influence the robustness of your results in unexpected ways.

- Take good notes of what match and what do not match in the various datasets you merge together. 
- Keep all raw data
- Keep a full dataset, which is a `full_join` of all.
- Keep a "smallest" dataset, usually the one you use to fit your full model

